---
title: "Generate Accumulated Cost Surfaces with Irregular Landscape Graphs"
author: "Joseph Stachelek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: irlgraph.bib
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Generate Accumulated Cost Surfaces with Irregular Landscape Graphs}
---

This vignette describes how to generate irregular landscape graphs in order to produce accumulated cost surfaces. Accumulated cost surfaces generated in this manner have several benefits as detailed by @etherington2012. Most notably, there is a large decrease in processing speed relative to regular graphs. There may be even larger decreases in processing speed relative to straight implementations of Dijkstra's algorithm (see @gdistance).

##Installation

Currently the `ilgraph` package can only be installed by compiling from source. One easy way to compile `R` packages is to use the `devtools` package [@devtools]. Eventually, `ilgraph` may be submitted to CRAN where compiled binaries would be made available ( [http://cran.r-project.org](http://cran.r-project.org)).

```{r echo=FALSE, eval=FALSE}
devtools::install_github("jsta/ilgraph")
```

##Generating Irregular Landscape Graphs

The @etherington2012 method of constructing irregular landscape graphs involves selecting a subset of points contained in the analysis domain. These subsets include:

1. Points of Interest

Optional points supplied to the `poicoords` argument of the `ir_graph` function.

2. Very important point cells

Cells that lie on either side of a boundary between different cost-categories in the underlying cost raster.

3. Landscape limit cells

Cells on the boundary of null data cells and the cost raster extent.

4. Ecological grain cells

A randomly selected number of cells. The proportion of cells selected in this manner is supplied to the `grainprop` argument of the `irl_graph` function.

5. Null data cells

&nbsp;

&nbsp;

The following code block loads the cost raster from @etherington2012 and calls the `irl_graph` function to generate an irregular landscape graph. Next, a sequence of plotting commands are called in order to visualize the point subsets in relation to the cost raster.

```{r eval=TRUE, fig.align='center', fig.height=4, fig.width=6}
library(irlgraph)
library(raster, quietly = TRUE)

set.seed(123) #make reproducible

dm <- as.matrix(read.delim(system.file("extdata/etherington20120code/cost-surface20x20.txt", package = "irlgraph"), skip = 6, na.strings = "-9999", header = FALSE, sep = " "))

costsurf <- raster::raster(nrows=dim(dm)[1],ncols=dim(dm)[2],resolution=1,xmn=0,
xmx = dim(dm)[1], ymn = 0, ymx = dim(dm)[2])
costsurf[] <- dm

par(mar = c(0,0,3,0))
plot(costsurf, box = FALSE, axes = FALSE)

poicoords <- matrix(c(10.5, 10.5), ncol = 2)
graph <- irl_graph(dm = dm, poicoords)

legend("top", inset  = c(0, -0.2), legend = c("Points of interest", "Very important",
"Landscape limit"), col = viridis::viridis(5)[1:3], pch = 1 , horiz = TRUE,
xpd = TRUE, cex = 0.7, bty = "n")

legend("top", inset  = c(0, -0.1), legend = c("Null data", "Ecological grain"),
col = viridis::viridis(5)[4:5], pch = 1, horiz = TRUE, xpd = TRUE, cex = 0.7, bty = "n")

points(xyFromCell(costsurf, graph$poicells), col = viridis::viridis(5)[1])
points(xyFromCell(costsurf, graph$vipcells), col = viridis::viridis(5)[2])
points(xyFromCell(costsurf, graph$limitcells), col = viridis::viridis(5)[3])
points(xyFromCell(costsurf, graph$nullcells), col = viridis::viridis(5)[4])
points(xyFromCell(costsurf, graph$graincells), col = viridis::viridis(5)[5])

```

The underlying graph can be visualized with the following call to the `plot` command. This preliminary graph includes edge connections to null cell nodes. Note that these nodes and any edge connections to these cells are stripped in the code underlying the `ir_graph` function before the final graph is constructed.

```{r eval=TRUE, fig.height=4, fig.width=6, fig.align='center'}
par(mar=c(2,0,0,0))
plot(graph$tri, wlines = "triang", xlab = "", ylab = "", showrect = FALSE, pch = 7)
```

##Generating Accumulated Cost Surfaces

Partial accumlated cost surfaces can be generated from an irregular landscape graph by specifying a starting node to the `snode` or `scoord` arguments of the `get_path` function. 

```{r eval=TRUE, fig.align='center', fig.height=4, fig.width=6}
result<-acc_path(graph$graph, scoord = c(10.5, 10.5), costsurf = costsurf)

par(mar=c(0,0,3,0))
plot(result, box = FALSE, axes = FALSE)
points(10.5, 10.5)
```

These partial surfaces can be converted to complete accumulated cost surfaces by using the `impute_na` function to perform nearest neighbor interpolation.

```{r eval=TRUE, fig.align='center', fig.height=4, fig.width=6}
result <- impute_na(costsurf, graph$cells, graph$nullcells, result, graph$cellcoords)

par(mar=c(0,0,3,0))
plot(result, box = FALSE, axes = FALSE)
points(10.5, 10.5)
```


#References